---
title: "Final Presentation"
subtitle: "Jack Fay"
format: revealjs
---
```{r}
library(tidyverse)
library(plotly)
library(kableExtra)

fifa_full <- read_csv("data/fifa_15_23.csv")

fifa_line1 <- fifa_full |> filter(short_name %in% c("L. Messi", "Cristiano Ronaldo"))|>
      group_by(short_name, fifa_version)|>
      summarise(overall = median(overall, na.rm = TRUE),
                  pace = median(pace, na.rm = TRUE),
                   shooting = median(shooting, na.rm = TRUE),
                   passing = median(passing, na.rm = TRUE),
                   dribbling = median(dribbling, na.rm = TRUE),
                   defending = median(defending, na.rm = TRUE),
                   physic = median(physic, na.rm = TRUE))
fifa_line2 <- fifa_full |> filter(short_name %in% c("L. Messi", "Cristiano Ronaldo"))|>
      group_by(short_name, fifa_version)|>
      summarise(overall = median(overall, na.rm = TRUE),
                  pace = median(pace, na.rm = TRUE),
                   shooting = median(shooting, na.rm = TRUE),
                   passing = median(passing, na.rm = TRUE),
                   dribbling = median(dribbling, na.rm = TRUE),
                   defending = median(defending, na.rm = TRUE),
                   physic = median(physic, na.rm = TRUE))
fifa_bar1 <- fifa_full|>
  filter(short_name %in% c("L. Messi", "Cristiano Ronaldo"), fifa_version == 15)|>
  pivot_longer(cols = 19:24, names_to = "stat", values_to = "value")|>
  group_by(short_name, stat)|>
  summarise(value = median(value), 
            overall = median(overall))
fifa_bar2 <- fifa_full|>
  filter(short_name %in% c("L. Messi", "Cristiano Ronaldo"), fifa_version == 23)|>
  pivot_longer(cols = 19:24, names_to = "stat", values_to = "value")|>
  group_by(short_name, stat)|>
  summarise(value = median(value), 
            overall = median(overall))
plot1 <- ggplot(data = fifa_line1,
           aes(x = fifa_version, y = shooting, color = short_name, label = overall)) +
      geom_point()+
      geom_line() +
      labs(title = "Shooting Comparison Over Time", x = "FIFA Version", color = "Player", y = "Shooting")
  
plot2 <- ggplot(data = fifa_line2,
           aes(x = fifa_version, y = pace, color = short_name, label = overall)) +
      geom_point()+
      geom_line() +
      labs(title = "Pace Comparison Over Time", x = "FIFA Version", color = "Player", y = "Pace")+
  scale_color_brewer(palette = "Set1")
plot3 <- ggplot(fifa_bar1, 
       aes(x = stat, y = value, group = short_name, fill = short_name, label = value))+
  geom_col(position = "dodge", color = "black")+
      labs(x = "Statistics", y = "FIFA 15", title = "2015 Comparison")
plot4 <- ggplot(fifa_bar2, 
       aes(x = stat, y = value, group = short_name, fill = short_name, label = value))+
  geom_col(position = "dodge", color = "black")+
      labs(x = "Statistics", y = "FIFA 23", title = "2023 Comparison")+
  scale_fill_brewer(palette = "Set1")


random_player_table <- fifa_full|>
  filter(short_name == "Neymar Jr", fifa_version == 22)|>
  select(6, 11, 12, 16:24)|>
  slice(1)
```
## Data 
* Kaggle Video Game Dataset
* Player data from Career Mode in FIFA 15 to FIFA 23
* Player Ratings, Performance Attributes, and Demographic Info
* Allows for multiple comparisons of same players across seasons
* ShinyApp subset only contains male players rated 80+


## Multi Season Static Visuals
<div class="scrollable-container" style="max-height: 500px; overflow-y: scroll;">
```{r}
ggplotly(plot1, tooltip = "label", height = 500)
```
```{r}
ggplotly(plot2, tooltip = "label", height = 500)
```
</div>
## Single Season Static Visuals
<div class="scrollable-container" style="max-height: 500px; overflow-y: scroll;">
```{r}
ggplotly(plot3, tooltip = "label", height = 500)
```
```{r}
ggplotly(plot4, tooltip = "label", height = 500)
```
</div>

## Random Player Game
```{r}
knitr::kable(random_player_table) |>
  kable_styling(full_width = TRUE, font_size = 16, bootstrap_options = "condensed")

```

* FIFA Version: 22
* Player Positions: LW, CAM
* Nationality: Brazil
* League Name: Ligue 1
* Club Name: Paris Saint Germain
* Player Name..................... Neymar Jr

## Future Work 
* Expand dataset to include female players 
* Expand dataset to include all overalls
* Examine which statistics influence overall most in each season
* Explore models to predict real life salary 
* Predict future card ratings based on past ratings and current performance




